---
title: 'Statistical Inference Course Project'
author: "Tobias Ford"
date: '`r Sys.Date()`'
output:
  html_document:
    fig.captions: yes
    keep_md: yes
    toc: no
---

```{r set-options, echo=FALSE}
options(width = 80, digits=2)
set.seed(42)
```

# Statistical Inference Course Project : Comparing Random Distributions

## Overview

Investingating the exponential distribution in R and comparing it to the Central Limit Theorem.
The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. 
The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. 

Set lambda = 0.2 for all of the simulations. 
You will investigate the distribution of averages of 40 exponentials. 
Note that you will need to do a thousand simulations.

```{r}
Lambda = 0.2
Exponentials = 40
Simulations = 1000
```

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should

1. Show the sample mean and compare it to the theoretical mean of the distribution.
2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.

In point 3, focus on the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials.

## Simulations

Let's start by doing the simulations
```{r}
SampleMeans <- sapply(1:Simulations, function(x) { mean(rexp(Exponentials, Lambda)) } )
```
The first ten samples are **`r head(SampleMeans,10)`**.

## Sample Mean vs. Theoretical Mean 

Determine mean of all the simulated means of the 40x random exponential distributions and then determine the diffference between Sample Mean and Theoretical Mean. Theoretcial Mean and Standard Deviation is 1/Lambda. 

```{r}
OverallSampleMean = mean(SampleMeans)
TheoreticalMean = 1/Lambda 
```
**Analysis** : The Sample Mean is **`r OverallSampleMean`**. The Theoretical Mean is **`r TheoreticalMean`**. Sample Mean is very close (**`r TheoreticalMean-OverallSampleMean`**) to Theoretical Mean. Close Enough!

Determine Diffference between Sample Standard Deviation and Theoretical Standard Deviation.

```{r}
OverallSampleStandardDeviation = sd(SampleMeans)
TheoreticalStandardDeviation =  1/Lambda / sqrt(40) 
```

**Analysis** : The Sample Standard Deviation is **`r OverallSampleStandardDeviation`**. The Theoretical Standard deviation is **`r TheoreticalStandardDeviation`**. Close Enough! Sample Standard Deviation is very close (**`r OverallSampleStandardDeviation - TheoreticalStandardDeviation`**) to the Theoretical Standard Deviation.

## Sample Variance vs. Theoretical Variance

Variance of the Simulations and Expected Variance 

```{r}
SampleVariance = var(SampleMeans)
ExpectedVariance = ( 1/Lambda / sqrt(40) )^2
```

**Analysis** : The Sample Variance is **`r SampleVariance`**. The Expected Variance is **`r ExpectedVariance`**. Close Enough! Sample Variance is very close (**`r SampleVariance - ExpectedVariance`**) to the Expected Variance.

## Distribution
Via figures and text, explain how one can tell the distribution is approximately normal.

```{r, echo=FALSE, fig.width=5, fig.height=2.5, fig.cap="**Fig.1 Distribution at 1000 samples of the means of 40 exponentials, overlayed with the normal distribution with the expected mean and standard deviation.**" }
library(ggplot2)
df = data.frame(means = SampleMeans)
ggplot(data = df, aes(x=means)) + 
	    geom_histogram(aes(y = ..density..), 
	    			 fill = I("blue"), 
				 binwidth = 0.2, 
				 color = I("darkblue")) + 
	    stat_function(fun = dnorm, arg = list(mean = TheoreticalMean, sd = TheoreticalStandardDeviation))+
	    geom_vline(xintercept=5, color=I("red"))
```

**Analysis**: The distribution of the sampled means fits into the normal distribution defined by TheoreticalMean (1/lambda) and TheoreticalStandardDeviation. Looks normal to me at 1000 samples. What would this look like for 10000? See Appendix

### Comparison to a large collection of random exponentials.

What is the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials?

```{r, echo=FALSE, fig.cap="**Fig.2 Distribution of 1000 exponentials.**", fig.width=6, fig.height=4 }
hist(rexp(1000, Lambda))
```

**Analysis**: As we've seen, 1000 Averages of 40 exponentials is normally distributed whereas a large collection (1000) of random exponentials fits closer to what an exponential curve looks like.

## Apendix

### What does a distribution of 10000 samples of the means of 40 exponentials look like?

```{r, echo=FALSE, fig.cap="**Fig.3 Distribution at 10000 samples of the means of 40 exponentials overlayed by the normal distribution.**" }
Simulations = 10000
SampleMeans <- sapply(1:Simulations, function(x) { mean(rexp(Exponentials, Lambda)) } )

df = data.frame(means = SampleMeans)
ggplot(data = df, aes(x=means)) + 
	    geom_histogram(aes(y = ..density..), 
	    			 fill = I("blue"), 
				 binwidth = 0.2, 
				 color = I("darkblue")) + 
	    stat_function(fun = dnorm, arg = list(mean = TheoreticalMean, sd = TheoreticalStandardDeviation))+
	    geom_vline(xintercept=5, color=I("red"))
```

**Analysis**: Even Closer to the normal distribution with more data.
