---
title: 'Statistical Inference Course Project'
author: "Tobias Ford"
date: '`r Sys.Date()`'
output:
  html_document:
    fig.captions: yes
    keep_md: yes
    toc: yes
  pdf_document:
    toc: yes
    latex_engine: xelatex
---

```{r set-options, echo=FALSE}
options(width = 80)
```

# Statistical Inference Course Project : Comparing Random Distributions

## Overview

Investingating the exponential distribution in R and comparing it to the Central Limit Theorem.

The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. 

The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. 

Set lambda = 0.2 for all of the simulations. 


```{r}
Lambda = 0.2
```

You will investigate the distribution of averages of 40 exponentials. 

```{r}
Exponentials = 40
```

Note that you will need to do a thousand simulations.

```{r}
Simulations = 1000
```

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should

1. Show the sample mean and compare it to the theoretical mean of the distribution.
2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.

In point 3, focus on the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials.

### Inital Setup
```{r}
set.seed(42)
```

### What does the exponential distribution look like again?

```{r}
x <- seq(0, 4, length=100)
hx <- dexp(x)
plot(x, hx, type='l')
```

## Simulations

Let's start by doing the simulations

```{r}
SampleMeans <- sapply(1:Simulations, function(x) { mean(rexp(Exponentials, Lambda)) } )

head(SampleMeans,10)
```

## Sample Mean vs. Theoretical Mean 

Include figures with titles; Highlight the means you are comparing
Include text that explains the figures and what is shown on them.. and provides appropriate numbers

Mean of all the simulated means of the 40x random exponential distributins 

```{r}
OverallSampleMean = mean(SampleMeans)
OverallSampleMean
```

Standard Deviation of all the simulated means of the 40x random exponential distributins 
```{r}
OverallSampleStandardDeviation = sd(SampleMeans)
OverallSampleStandardDeviation
```

Theoretcial Mean and Standard Deviation is 1/Lambda

```{r}
TheoreticalMean = 1/Lambda 
OverallSampleMean - TheoreticalMean
```
Analysis : Close Enough!

```{r}
TheoreticalStandardDeviation = ( 1/Lambda / sqrt(40) )
OverallSampleStandardDeviation - TheoreticalStandardDeviation
```
Analysis : Close Enough!


## Sample Variance vs. Theoretical Variance

 Include figures (output from R) with titles. Highlight the variances you are comparing. Include text that explains your understanding of the differences of the variances.

Variance of the Simulations

```{r}
SampleVariance = var(SampleMeans)
SampleVariance
```

Expected Variance 
```{r}
ExpectedVariance = ( 1/Lambda / sqrt(40) )^2
ExpectedVariance
```

Difference between the two..
```{r}
ExpectedVariance - SampleVariance
```
Analysis : Close Enough!

## Distribution
Via figures and text, explain how one can tell the distribution is approximately normal.

```{r}
library(ggplot2)
df = data.frame(means = SampleMeans)
ggplot(data = df, aes(x=means)) + 
	    geom_histogram(aes(y = ..density..), 
	    			 fill = I("blue"), 
				 binwidth = 0.2, 
				 color = I("darkblue")) + 
	    stat_function(fun = dnorm, arg = list(mean = TheoreticalMean, sd = TheoreticalStandardDeviation))+
	    geom_vline(xintercept=5, color=I("red"))
```

The distribution of the sampled means fits into the normal distribution defined by TheoreticalMean (1/lambda) and TheoreticalStandardDeviation.
Analysis: Looks normal to me at 1000 samples. What would this look like for 10000?

```{r}
Simulations = 10000
SampleMeans <- sapply(1:Simulations, function(x) { mean(rexp(Exponentials, Lambda)) } )

df = data.frame(means = SampleMeans)
ggplot(data = df, aes(x=means)) + 
	    geom_histogram(aes(y = ..density..), 
	    			 fill = I("blue"), 
				 binwidth = 0.2, 
				 color = I("darkblue")) + 
	    stat_function(fun = dnorm, arg = list(mean = TheoreticalMean, sd = TheoreticalStandardDeviation))+
	    geom_vline(xintercept=5, color=I("red"))
```
Analysis: Even Closer to the normal distribution with more data.