---
title: 'Statistical Inference Course Project : Tooth Growth'
author: "Tobias Ford"
date: '`r Sys.Date()`'
output:
  html_document:
    fig.captions: yes
    keep_md: yes
    toc: no
---

```{r set-options, echo=FALSE}
options(width = 80, digits=2)
set.seed(42)
library(knitr)
library(ggplot2)
```
## Overview

Now in the second portion of the assignment, we're going to analyze the ToothGrowth data in the R datasets package.

The ToothGrowth is a data set demonstrating a response in the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods / supplement types either orange juice or ascorbic acid (a form of vitamin C and coded in the data as VC).

[Link to Explanation of Tooth Growth data](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/ToothGrowth.html)

## Load the ToothGrowth data and perform some basic exploratory data analyses

```{r, echo=FALSE}
data(ToothGrowth)
```

First 3 rows of the ToothGrowth data set and the header.

```{r,echo=FALSE}
kable(head(ToothGrowth, 3))
```

```{r, echo=FALSE}
ToothGrowth$dose<-as.factor(ToothGrowth$dose)
```

```{r}
summary(ToothGrowth)
```

## Data Analysis

First let's visualize the data to gain some intuition about what is going on in it. Graph the data split amonst the different supplement types.

```{r, echo=FALSE, fig.width=4, fig.height=3}
ggplot(data=ToothGrowth, aes(x=supp, y=len)) + 
			 geom_boxplot(aes(fill=supp))+ 
			 xlab("Supplement type") +ylab("Tooth length")
```

From the plot, it's pretty hard to see how supplement type positively impacted tooth growth.
Even thought the 'OJ' supplment type shows a higher density then 'VC', 'VC's range of tooth length encompses 'OC'.

Now check the effect of vitamin C dose on tooth length.

```{r, echo=FALSE, fig.width=4, fig.height=3}
ggplot(data=ToothGrowth, aes(x=dose, y=len)) + 
			 geom_boxplot(aes(fill=dose))+ 
			 xlab("Dose in mg") +ylab("Tooth length")
```

From the plot, it's pretty easy to see how dosage size positively impacted tooth growth.
Let's use what we've learned in class to show this with statistical inference.

### Conclusions via Statistical Inference

The three general questions to answer.. 

Are the tooth length of the guinea pigs impacted positively by a supplement type / delivery methods? 
Are the tooth length of the guinea pigs impacted positively by a supplement type / delivery methods at certain dosage amounts?
Are the tooth length of the guinea pigs impacted positively by changing dosage amounts?

A t test will be performed on each of the potential hypothesis which caused tooth length to increase.
Recall.. no impact on tooth length is the null hypothesis and the alternative hypothesis is that a test variation caused positive tooth length.

```{r}
Length <- ToothGrowth$len; SupplementType <- ToothGrowth$supp; Dose <- ToothGrowth$dose

tOJtoVC <- t.test(len ~ supp, ToothGrowth)
tSTat0.5 <- t.test(len ~ supp, ToothGrowth[Dose == 0.5, ])
tSTat1 <- t.test(len ~ supp, ToothGrowth[Dose == 1, ])
tSTat2 <- t.test(len ~ supp, ToothGrowth[Dose == 2, ])
t1to2 <- t.test(Length[Dose==2], Length[Dose==1], var.equal = TRUE)
t0.5to1 <- t.test(Length[Dose==1], Length[Dose==0.5], var.equal = TRUE)
t0.5to2 <- t.test(Length[Dose==2], Length[Dose==0.5], var.equal = TRUE)
```

```{r, echo=FALSE}
df <- data.frame(
   'null hypotheses'=c('OJ to VC Delivery Method',
   	 'OJ to VC at 0.5mg Dosage',
   	 'OJ to VC at 1mg Dosage',
   	 'OJ to VC at 2mg Dosage',
   	 '1mg to 2mg Dosage',
	 '0.5mg to 1mg Dosage',
	 '0.5mg to 2mg Dosage'),
   'p value'=c(tOJtoVC$p.value, 
      tSTat0.5$p.value, 
      tSTat1$p.value, 
      tSTat2$p.value, 
      t1to2$p.value, 
      t0.5to1$p.value, 
      t0.5to2$p.value),
   conf.int.low=c(
	tOJtoVC$conf.int[1], 
	tSTat0.5$conf.int[1],
	tSTat1$conf.int[1],
        tSTat2$conf.int[1],
	t1to2$conf.int[1], 
	t0.5to1$conf.int[1], 
	t0.5to2$conf.int[1]),
   conf.int.high=c(
	tOJtoVC$conf.int[2], 
	tSTat0.5$conf.int[2],
	tSTat1$conf.int[2],
        tSTat2$conf.int[2],
	t1to2$conf.int[2], 
	t0.5to1$conf.int[2], 
	t0.5to2$conf.int[2]),
   is.zero.within.conf.int=c(
	tOJtoVC$conf.int[1] < 0 && 0 < tOJtoVC$conf.int[2], 
	tSTat0.5$conf.int[1] < 0 && 0 < tSTat0.5$conf.int[2], 
	tSTat1$conf.int[1] < 0 && 0 < tSTat1$conf.int[2], 
	tSTat2$conf.int[1] < 0 && 0 < tSTat2$conf.int[2], 
	t1to2$conf.int[1] < 0 && 0 < t1to2$conf.int[2], 
        t0.5to1$conf.int[1] < 0 && 0 < t0.5to1$conf.int[2], 
	t0.5to2$conf.int[1] < 0 && 0 < t0.5to2$conf.int[2]
     )
)
```

```{r, echo=FALSE}
kable(df)
```

The p-value of the t test for the change in supplement type is around 0.06, very close to the default significance level of 0.05, and the confidence interval includes 0. This is evidence we can't definitively reject the null hypothesis of a non impactful comparison of supplement type against positive effect on tooth length. After this analysis we can't definitively confirm or deny that supplement type has a positive effect on tooth length overall.

The p-values of the t test for each specific dosage level versus which delivery method was used showed that for 0.5 and 1 mg dosages the null hypothesis was rejected, the p-values were very low and 0 wasn't in the confidence interval, suggesting for those two dosages that OJ improved tooth length. Not the case for 2mg.. the p-value is very clear in the opposite.. there was no impact on tooth length at 2mg for either delivery method.

The p-value of the t test for all three dosage changes is essentially 0 and the confidence interval in each case doesn't contain 0. Evidence that we can reject the null hypothesis in comparison of dosages versus tooth length. Therefore, we can assume that a increase in dosage creates an positive effect on tooth length. 

